---
title: "Data processing documentation"
author: "Byron Jaeger"
date: "February 7, 2019"
output:
  pdf_document: default
  html_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

from_cran <- c("haven","magrittr","dplyr","purrr","diagram")

# check the installs:
for (pkg in from_cran)
  library(pkg, character.only = TRUE)

```

# CARDIA data processing file

This file contains the following: 

* Source codes used to merge CARDIA data, documented with my comments and placed in code blocks interspersed throughout the document. 
* Source codes used to make exclusions, accompanied by a step-by-step description of the exclusions (see final section)
* Source codes used to derive variables along with my comments. 
* plain text that accompanies this code and describes its purpose. 

This file is also reproducible using Rstudio and Rmarkdown. Steps: 

1. Create a new R project directory
2. Save datafiles in this directory within a folder that is named "Datasets"
3. Open the accompanying .Rmd file.
4. Run the following setup code:

```{r, eval=FALSE}

# Install and load packages -----------------------------------------------

# If you have any issues with the installs, please contact me at:
# bcjaeger@uab.edu

from_cran <- c("haven","magrittr","dplyr","purrr","diagram")

install.packages(from_cran, repos = "http://cran.rstudio.com")

# check the installs:
for (pkg in from_cran)
  library(pkg, character.only = TRUE)


```
5. Run the accompanying .Rmd file.
6. If you have any trouble, contact me at bcjaeger@uab.edu

# CARDIA data processing

## Step 1: read in the data

```{r}

data=cardia=list()

# Read files from CARDIA ancillary ABPM study
# These paths will not work on your computer, but the files 
# should be available in the CARDIA database
schwartz_abpm_loc=file.path(
  'O:','REGARDS','CARDIA','Data','Population','ABPM_Schwartz'
)

uab_slp=read_sas(
  file.path(
    schwartz_abpm_loc,'final_summary_dataset_uab_v1.sas7bdat'
  )
)

nw_slp<-read_sas(
  file.path(
    schwartz_abpm_loc,'final_summary_dataset_nw_v1.sas7bdat'
  )
)

cardia$slp=rbind(uab_slp,nw_slp) 

```

## Step 2: Outline the exclusion steps

```{r}

# initialize exclusion info
nobs=labs=list()

nobs[[1]]=5115
labs[[1]]="CARDIA participants"

nobs[[2]]=nrow(cardia$slp)
labs[[2]]="Participants who underwent \n24-hour ABPM"

# remove participants with insufficient sleep readings

cardia$slp%<>%
  dplyr::filter(GE_5_Valid_sleep==1,
                GE_10_Valid_awake==1) 

nobs[[3]]=nrow(cardia$slp)
labs[[3]]="Participants with \u2265 10 awake \nand \u2265 5 asleep ABPM readings"

```

## Step 3: Apply data processing steps

```{r}

# some variable derivations for BP during sleep
cardia$slp %<>%
  dplyr::select(id,
                slp_sbp=mn_sysbp_SL,
                pct_dip_sbp=Pct_sysbp_dip,
                slp_dbp=mn_diabp_SL,
                slp_dur=sleep_duration) %>%
  dplyr::mutate(nhtn=ifelse(slp_dbp>=70|slp_sbp>=120,1,0),
                ndbp=ifelse(pct_dip_sbp<10, 1,0),
                slp_dur=as.numeric(slp_dur)/60^2)

exclusion_cascade=list(cardia=list(labs=labs,nobs=nobs))

# read in ABPM forms
cardia$abpm_forms <- 
  read_sas(file.path('O:','REGARDS','JHS','Jaeger',
                     'Miscellaneous','Swati',
                     'cardia_abpm_781_updated.sas7bdat')) %>% 
  set_names(tolower(names(.))) %>%
  dplyr::select(id=short_id,age,neck_cm,htn_meds)

# Read in data from my CARDIA data request
cardia$data_forms <- 
  read_sas(file.path("O:","REGARDS","CARDIA","Data","Population",
                     "A1808 REQ 4 11 2018 (Byron Jaeger)",
                     "a1808req4_11_2018.sas7bdat")) %>%
  magrittr::set_names(tolower(names(.))) %>% 
  dplyr::filter(short_id%in%cardia$slp$id) %>% 
  # variable derivations
  dplyr::mutate(
    race=factor(ifelse(a01race1==5,'White','Black')),
    sex=factor(ifelse(a01sex==1,'Male','Female')),
    alcohol=factor(ifelse(i07drink==1,'None','Some')),
    smoke=factor(i10smoke,levels=c(0,1,2),
                 labels=c("Never","Former","Current")),
    diabetes=ifelse(i08diab==1,0,
                    ifelse(i08diab==2,1,NA)),
    acr=ifelse(il1ratio>30, 1, 0))%>%
  # variable selection and re-naming
  dplyr::select(id=short_id,
                sex=sex,
                race=race,
                educ=i03ed,
                alcohol,
                smoke,
                height_cm=i20hgt,
                weight_lbs=i20wgt,
                waist_cm=i20wst,
                cln_sbp=i02savg,
                cln_dbp=i02davg,
                ur_cre=il1ucreat,
                ur_alb=il1ualb,
                acr=acr,
                egfr=il7gfr,
                hba1c=i0yhba1c,
                ldl=il1ldl,
                hdl=il1hdl,
                chol=il1chol,
                glucose=il7glu,
                diabetes) %>%
  mutate(ur_alb=ur_alb*1000)

# setting some very questionnable neck circumference values to NA
cardia$abpm_forms$neck_cm[cardia$abpm_forms$neck_cm<20]=NA
cardia$abpm_forms$neck_cm[cardia$abpm_forms$neck_cm>400]=NA

# Merge all of the data by id
data$cardia=reduce(cardia,left_join,by='id')%>%dplyr::select(-id)%>%
  mutate(hispanic=0,race=as.character(race))

glimpse(data$cardia)

```

## Step 4: Print exclusion chart

```{r, fig.width=8, fig.height=10}

plot_exclusion_cascade <- function(nobs, labs){
  
  par(mar = c(1, 1, 1, 1))
  nrows=length(nobs)
  openplotmat()
  pos<-coordinates(rep(1,nrows))
  
  for(i in 1:nrows){
    
    if(i<nrows){
      Arrows(x0=pos[i,1],y0=pos[i,2]-1/(nrows*3.8),
             x1=pos[i+1,1],y1=pos[i+1,2]+1/(nrows*3.8),lwd=2)
    }
    textrect(mid = pos[i,], radx = 1/2, rady = 1/(nrows*4),
             lab = paste0(labs[[i]],'\n N = ',format(nobs[[i]],big.mark=',')), 
             box.col='white', col='black',
             cex = 2, shadow.size = 0.001)
    
  }
}

plot_exclusion_cascade(nobs, labs)

```


# Computing environment

Please find information about the computing environment used to create this file below:

  * File creation date: `r Sys.Date()`
  * `r R.version.string`
  * R version (short form): `r getRversion()`
  * `mosaic` package version: `r packageVersion("mosaic")`
  * Additional session information

```{r echo=FALSE}
sessionInfo()  # could use devtools::session_info() if you prefer that
```
  